---
title: "Lab 9 Part 1 - Survival analysis"
author: "Isabel Sassoon"
output:
  pdf_document: default
  html_notebook: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Survival Analysis

This notebook walks through the analysis of the lung data from the survival library.

```{r}
library(survival)
```

The lung data has many columns

```{r}
help("lung")
```

In order to focus I subset the data into a smaller set of columns:

```{r}
lung.analysis<-subset(lung, select = c("time", "status", "sex"))
```


## To estimate the survival curve using KM

```{r}
s.survfit <- survfit(Surv(lung.analysis$time, lung.analysis$status)~ 1, data=lung.analysis )
s.survfit
```

and to plot it:

```{r}
plot(s.survfit, mark.time = TRUE, conf.int = FALSE)
title(main="Survival Curve for Lung data", xlab="Days", ylab = "Proportion of Patients still alive")
```

## Research Question: Does Gender affect survival?

Here we use survdiff to see if there is a difference.

```{r}
lung.gender.survfit<-survfit(Surv(lung.analysis$time, lung.analysis$status)~ lung.analysis$sex, data=lung.analysis)

lung.gender.survdiff<-survdiff(Surv(lung.analysis$time, lung.analysis$status)~ lung.analysis$sex, data=lung.analysis)

lung.gender.survdiff

plot(lung.gender.survfit, mark.time = TRUE, col=c("black", "grey75"), lty=1:2)
title(main="Survival Curve by Gender", xlab="Days", ylab = "Proportion of Patients still alive")
legend("topright", c("1- Male", "2-Female"), lty=1:2, col=c("black", "grey75"))

```
The p-value is significant and this supports what we see in the survival curves. There is a difference in survival curves between Male and Female patients.

# Another possible research question

```{r}
lung.perf.survfit<-survfit(Surv(lung$time, lung$status)~ lung$ph.ecog, data=lung)

lung.perf.survdiff<-survdiff(Surv(lung$time, lung$status)~ lung$ph.ecog, data=lung)

lung.perf.survdiff

plot(lung.perf.survfit, mark.time = TRUE, col=c("black", "grey75"), lty=1:4)
title(main="Survival Curve by ECOG performance score", xlab="Days", ylab = "Proportion of Patients still alive")
legend("topright", c("0", "1", "2", "3"), lty=1:4, col=c("black", "grey75"))
```

We can see that there is one category with only one case. Perhaps this can be removed or grouped with group ecog=2?


## Cox Proportional Hazards (OPTIONAL)

```{r}
lung.gender.ph<-coxph(Surv(lung.analysis$time, lung.analysis$status)~ lung.analysis$sex, data=lung.analysis)
summary(lung.gender.ph)

```

The exp(coef) column contains $exp(\beta_1)$
This is the hazard ratio – the multiplicative effect of that variable on the hazard rate (for each unit increase in that variable). So, for a categorical variable like gender (in this case), going from male (baseline) to female results in approximately ~40% reduction in hazard.
Recall that the CoxPH model is a linear model of the natural log of the hazard at time t, denoted $h(t)$, as a function of the baseline hazard $(h_0(t))$

$log(h(t))=log(h_0(t))+ \beta_1x_1 + \dots + \beta_px_p$

if both sides are exponentitated:

$h_1(t)=h_0(t) \times exp(\beta_1x_1)$

Rearranging makes it possible to estimate the hazard ratio:

$HR(t)=\frac{h_1(t)}{h_0(t)}=exp^{\beta_1}$

