---
title: "Lab 8 Part 1"
author: Dr. Isabel Sassoon
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---


# Table Analysis

In this lab we will explore how to perform table analysis. This is useful when there are multiple columns of data that are categorical.

```{r}
# installing a new library
library(vcdExtra)
```

The data we are using is called Arthritis

```{r}
help(Arthritis)
```


Lets explore the data

```{r}
summary(Arthritis)
```


We can see there is an ID, Gender, Age and whether the person had a Marked Improvement, Some improvement or none.

Table analysis can help us if we want to find out if there is a relationship between the improvement and the treatment (for example).

```{r}
table(Arthritis$Improved, Arthritis$Treatment)
```



***Null hypothesis***: $H_0$, The treatment and improvement are independent
***Alternative hypothesis***: $H_1$, there is a relationship between the treatment and the improvement

```{r}
chisq.test(table(Arthritis$Improved, Arthritis$Treatment))
```


We can see that our $\chi^2$ is significant - so there is a dependence between treatment and outcome.

Note that there is no cell in the table that is so small that we need to consider using Fisher's exact test.

***

We can now take this one step further and see if we can answer the question: What attributes affect the success of the treatment?

If we wanted to model this relationship further we could transform the dependent variable from one with three values to one with two and then use Logistic regression:

Let's define the new dependent variable:

- 0 if there is none 
- 1. When there is an improvement (some or marked) 
 
```{r}
Arthritis$Improved.Ind<-ifelse(Arthritis$Improved=="None", 0,1)
```
 
Now we can model this:

```{r}
Arthritis.lr<-glm(Arthritis$Improved.Ind~Arthritis$Treatment, family = "binomial")
summary(Arthritis.lr)
```

This confirms our table analysis. We can also look at the coefficients as Odds Ratios

```{r}
exp(coef(Arthritis.lr))
```



But what about the other possible covariates?

```{r}
Arthritis2.lr<-glm(Arthritis$Improved.Ind~Arthritis$Treatment+ Arthritis$Sex+ Arthritis$Age, family = "binomial")
summary(Arthritis2.lr)
```

```{r}
exp(coef(Arthritis2.lr))
```

We can also start with the interactions model

```{r}
Arthritis3.lr<-glm(Arthritis$Improved.Ind~Arthritis$Treatment*Arthritis$Sex*Arthritis$Age, family = "binomial")
summary(Arthritis3.lr)
```

Note that all the interactions are not significant. The AIC is 97.18, which is very close to the one obtained without the interactions (100.06).