---
title: "QDA Lab 2 - Inferential Statistics"
output: html_notebook
---

```{r}
#Loading the required library
library(ggplot2)
```



In this lab we will be covering the following topics:

 -  Probability distributions
 -  confidence intervals

# Probability functions

There are four basic probability functions for each probability distribution in R. These begin with the same four prefixes: d, p, q or r followed by the root name of the probability that indentifies the specific distribution. For example for the **normal** distribution the name to use is **norm**.

 - d is for *”density”* and the corresponding function returns the value from the probability density function (continuous) or probability mass function (discrete).
 - p is for *”probability”* and the corresponding function returns a value from the cumulative distribution function.
 - q is for *”quantile”* and the corresponding function returns a value from the inverse cumulative distribution function.
 - r is for *”random"* and the corresponding function returns a value drawn randomly from the given distribution
 
 
## The normal distribution
 
When we need to work with different functions related to the normal distribution then the following syntax is used

To obtain the **density** of the normal distribution:

d+ name of the distribution (the value on the x axis)

So for example if we wanted to know the value of y (from the y axis), from the normal distribution curve that corresponds to x=2 then we would use:

```{r}
dnorm(2)
```

 
 If we wanted the value of y, but using a normal distribution with different means and s.d these are optional arguments in the function:
 
```{r}
dnorm(11,10,0.5)
```
 
*pnorm* is the cumulative distribution function for the normal distribution. By definition $pnorm(x) = P(X \leq x)$ and is the area under the normal density curve to the left of x. 

```{r}
pnorm(2)
```


The value above shows pnorm(2), the area under the normal density curve to the left of $x = 2$. From this we can see that the probability that a standard normal random variable takes on a value less than or equal to 2 is 0.977

*qnorm* is the quantile function of the standard normal distribution. If $qnorm(x)= k$ then k is the value such that $P(X \leq k) = x$. *qnorm* is the inverse function for *pnorm*. 

```{r}
qnorm(0.977)
```


From the chunk of code above  we have: qnorm(0.977) = qnorm(pnorm(2)) = 2.

*rnorm* generates random values from a standard normal distribution. The required argument is a number specifying the number of normal variables to produce. 

```{r}
x<-rnorm(20)
x1<-rnorm(2000)
```


The figure below illustrates rnorm(20) and rnorm(2000) the larger the number in the bracket the more "bell curved" the plot is.

```{r}
# use this space to graph x
hist(x)
hist(x1)
```

***

# Examples using other distributions

## Poisson Distribution

```{r}

x1=0:100
y2<-dpois(x1,lambda=50)

plot(x1,y2, type="h", main = "Poisson Distribution (Lambda=50)", ylab="Frequency", xlab="Count")
```


## Binomial Distribution

```{r}
y<-dbinom(x, size=10, prob = 0.5)

plot(x,y,type='h',
     main='Binomial Distribution (n=10, p=0.5)',
     ylab='Probability',
     xlab ='# heads',
     lwd=3)
```

## Uniform Distribution

```{r}
xu<-runif(100000, min=1, max=2)
xu<-as.data.frame(xu)
ggplot(xu, aes(x=xu)) + geom_density() + theme_classic() +
  labs(main="A Uniform Distribution", xlab="value", ylab="Density") + xlim(0,3)
```


## Exponential Distribution


```{r}
#exponential distribution

xe<-rexp(10000, rate = 10)
xe<-as.data.frame(xe)

ggplot(xe, aes(x=xe)) + geom_density() + theme_classic() +
  labs(main="An Exponential Distribution", xlab="value", ylab="Density") + xlim(0,3)
```


***

# Confidence Intervals in R

## Reading in from csv - use the skew data.csv

```{r}
data<-read.csv("data/skewdata.csv")
```

## Visualise the data

```{r}
summary(data)
```

This data contains only one variable. We can see its mean is 30.969.

## How to run confidence intervals for this variable - Z and T at the 95% confidence level

```{r}
mean.values<-mean(data$values)
sd.values<-sd(data$values)
```

In order to compute the confidence interval for a mean using Z we would apply this equation

: $$  (\bar{x} -  Z \times \frac{s}{\sqrt{n}}, \bar{x} + Z \times \frac{s}{\sqrt{n}}) $$

We have the value for $\bar{x}$ and s from the code chunk above.
We can get the value of Z 0.95 from R 



```{r}
qnorm(0.975)
```

Recall 0.95 confidence for a two sided test will result in us looking for Z 0.975 on each side. The number is the same and it is 1.96

Then we just need to combine this to compute the intervals

```{r}
ucl<-mean.values+qnorm(0.975)*sd.values/sqrt(30)
lcl<-mean.values-qnorm(0.975)*sd.values/sqrt(30)

```

```{r}
ucl
lcl
```

# If we replace Z with T

```{r}
qt(0.975, 29)
```

```{r}
uclt<-mean.values+qt(0.975,30)*sd.values/sqrt(30)
lclt<-mean.values-qt(0.975,30)*sd.values/sqrt(30)
```

```{r}
uclt
lclt
```

The confidence levels are very similar to the ones obtained with Z.

***
# OPTIONAL
## Confidence Intervals using Bootstrap

```{r}
data<-read.csv("data/skewdata.csv")

plot(c(0,30), c(0,60), type = "n", xlab = "Sample Size", ylab = "Confidence Interval")
for (k in seq(5,30,3)) {
  a<-numeric(10000)
  for (i in 1:10000){
    a[i]<-mean(sample(data$values,k,replace=T))
  }
  points(c(k,k), quantile(a,c(.025,.975)), type="b", pch=21, bg="red")
}
# from page 62 Crawley
```

# computing confidence intervals for the data used in lecture 2

The rivets strength data

```{r}
riv<-c(4.01, 4.28, 4.43, 4.04, 3.74, 4.43, 3.89, 4.73, 3.68, 4.55, 4.10, 3.77)

```

## add an example that includes:
generating a series of binary values
computing the confidence interval for the proportion

 