---
title: "Lab 6 Part 1"
output:
  html_document:
    df_print: paged
---

```{r}
library(ggplot2)
library(ggpubr)
library(xtable)
library(mgcv)
library(tree)

```

This covers multiple regression and focuses on approaches to model selection.

We are using a data set (seen in the lecture) on air pollution studies.

**How is ozone concentration related to wind speed, air temperature, and the intensity of solar radiation?**

In this case we have a *numerical dependent variable and all the explanatory variables are also numeric*.

```{r}
ozone.pollution<-read.csv("data/ozone.data.csv")
```

Lets explore the data:

```{r}
summary(ozone.pollution)
```

There are no missing values, extreme values, or strange patterns in the summaries.

We can explore the data graphically:

```{r}
pairs(ozone.pollution, panel = panel.smooth)
plot(ozone.pollution)
```


The response variable, ozone concentration is shown on the y-axis of the bottom row of these panels. There appears to be a strong negative relationship with wind speed, a positive correlation with temperature and an unclear relationship with radiation.

***

## OPTIONAL
One way to start a multiple regression problem is using a non-parametric smoother in a GAM.
```{r}
ozone.gam<-gam(ozone.pollution$ozone~s(ozone.pollution$rad)+s(ozone.pollution$temp)+s(ozone.pollution$wind))
summary(ozone.gam)
plot(ozone.gam)
```

This is useful as it can highlight some evident curvature in the relationships between the target and explanatory variables. 


Another optional approach to see if there are interactions, is using a regression tree.

```{r}
ozone.tree<-tree(ozone.pollution$ozone~., data = ozone.pollution)
plot(ozone.tree)
text(ozone.tree)
```

This can tell us about the structure of the data, and the top factors affecting the value of our dependent variable.

In this case it shows that temperature is the most important factor affecting ozone concentration (expressed as longer tree branches, which represent greater deviance).
Wind speed is important at both high and low temperature.
Radiation shows an interesting effect, although subtle where at low temperatures radiation matters at relatively low wind speeds. Whereas at high temperatures radiation matters at low wind speeds.

END OF OPTIONAL SECTION
***

From the plots of the data (and confirmed by the GAM and Tree - if using) there is reason to assume that some polynomial relation is possible. Therefore our **most complicated model** will include some quadratic terms and interactions.

```{r}
m1.lm<-lm(ozone~temp*wind*rad+I(temp^2)+ I(wind^2)+ I(rad^2), data=ozone.pollution )
summary(m1.lm)
```

This model has a significant F score, a great $r^2$ and some coefficients (but not all) are significant.

The three way interaction is clearly not significant so we can start to simplify by removing it.

```{r}
m2.lm<-lm(ozone~temp*wind*rad+I(temp^2)+ I(wind^2)+ I(rad^2)-temp:wind:rad, data=ozone.pollution)
summary(m2.lm)
```

We will now remove the least significant two way interaction

```{r}
m3.lm<-update(m2.lm,~.-wind:rad)
summary(m3.lm)
```

And now we remove the wind and temperature interaction

```{r}
m4.lm<-update(m3.lm, ~.-temp:wind)
summary(m4.lm)
```

Lets keep the marginally significant interaction between temp and rad, but leave out all the others. Also the least significant coefficient was the one for $rad^2$ so we remove it from our next model.

We can use the **update* function to stop too much typing!

```{r}
m5.lm<-update(m4.lm, ~.-I(rad^2))
summary(m5.lm)
```

Now we find that temp:rad is not significant so we will remove it from the next model

```{r}
m6.lm<-update(m5.lm, ~.-temp:rad)
summary(m6.lm)
```

This is looking promising, as all the coefficients are significant we don't need to remove any. But we should check the residuals

```{r}
plot(m6.lm)
```

We can see in the first plot that the spread of the residuals is greater for higher values on the x-axis. This is called heteroscedasticity. 
There is also a S shape emerging in the qq-plot.

In order to adjust for this a transformation of the dependent variable can help. In this case a log transformation. This is ok to use as long as there are no zeros in the dependent variable. 

Now once the dependent variable is transformed, the whole structure and linearity of the relationships changes. We will start from the beginning, but we can use the step function as a shortcut.

So firstly we go back to the full model or most complicated model, but with the transformed dependent variable:

```{r}
#log transform
m7.lm<-lm(log(ozone)~temp*wind*rad+I(temp^2)+ I(wind^2)+ I(rad^2),data=ozone.pollution)
summary(m7.lm)
```


Then use the step function to get to our *minimal adequate model*

```{r}
m8.lm<-step(m7.lm)
summary(m8.lm)
```

The step function creates a long output so what we are looking for is the last part, which proposes the *minimal adequate model* where the step wise process stops.

This model is:

$$log(ozone)=0.77 + 0.041 \times temp - 0.22 \times wind +0.0075 \times rad + 0.0074 \times wind^2 -0.000147 \times rad^2  $$

and check its residuals are obtained using:

```{r}
plot(m8.lm)
```

Which are looking much better!
