---
title: "QDA Lab 8 part 2 - a Solution"
author: "Isabel Sassoon"
date: "November 2020"
output:
  pdf_document: default
  html_notebook: default
---

# Part 1 - ARM data

```{r}
arm.folding<-arm
```

This data contains two columns only

```{r}
summary(arm.folding)
```
The two categorical variables are the gender of the participant and the arm crossed on top.

```{r}
table(arm.folding$gender, arm.folding$armcross)
```

We can test for independence using $\chi^2$, as the expected number in the cells is more than 5. Note that if that were not the case R would complain!

```{r}
arm.table<-table(arm.folding$gender, arm.folding$armcross)
chisq.test(arm.table)
```

We can see from here that this is not significant.

*** 
If you wanted to run Fisher's exact:

```{r}
fisher.test(arm.table)
```

Note: it is not necessary in this case but wanted to show that it would return the similar conclusion.
***

# Part 2 - back to titanic

```{r}
titanic<-titanic.all
```

Recall from week 7 - we need to make sure that the data is read in correctly:

```{r}
summary(titanic)
```

We need to make sure that both Survived and Pclass are treated as categorical, so I use as.factor to make sure that is the case.

```{r}
titanic$Survived<-as.factor(titanic$Survived)
titanic$Pclass<-as.factor(titanic$Pclass)
```


Lets look at table analysis for Survived vs Pclass

Are survival status and Passenger class independent? THat is the $H_0$ that we are testing here

```{r}
table(titanic$Survived, titanic$Pclass)
```

Lets use Chisq

```{r}
chisq.test(table(titanic$Survived, titanic$Pclass))
```

We can see that there is a significant relationship between Passenger travel class and survival status. In other words the probability of surviviing depends on the passenger's travel class.

If we wanted to model this probability then this is when logistic regression can help - and this is what we did in Lab 7 part 2.

***

## OPTIONAL: Using the gmodels library:

The Gmodels library is useful if you are used to SAS PROC means or PROC summary.

```{r}
library(gmodels)
```


```{r}
CrossTable(titanic$Survived, titanic$Pclass, digits=2, prop.r = TRUE, prop.c = TRUE,  prop.chisq = FALSE, 
           chisq = TRUE, fisher = TRUE)
```

As you can see it provides both tests. When you are reporting your findings you should still comment on the one that is appropriate to the situation.


