---
title: "CS5801 Coursework Template Proforma"
author: "student-id"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
version: 1.0
---

# 0. Instructions 

*1. Remove the (italicised) guidance text but keep the section headings.*  
*2. Add as many chunks of R code as required.*  
*3. Add descriptions of your analysis plans and explanations of your code and findings.  Please be detailed and where you have made choices explain the rationale for them.*  
*4. Write your report using RMarkdown.  For guidance see a [helpful blog](https://www.dataquest.io/blog/r-markdown-guide-cheatsheet/#tve-jump-17333da0719) or use the R Markdown cheatsheet which can be accessed from within RStudio by selecting `Help > Cheatsheets > R Markdown Cheat Sheet`.*  
*5. Your report should be clearly and professionally presented with appropriate use of cited external sources. (5 marks)*  
*6. It should also be easy to understand, with well-documented code following the principles of literate programming. (5 marks)*  

```{r}
# Add code here to load any required libraries with `library()`.  
# We suggest you use `install.package()` for any required packages externally to this document 
# since installation only need be done once.

install.packages("ggplot2")
install.packages("dplyr")
install.packages("dbplyr")
install.packages("foreign")
install.packages("ggpubr")
install.packages("xtable")
install.packages("mgcv")
install.packages("tree")
install.packages("validate")
install.packages("ggpubr")
install.packages("lubridate")
install.packages("GGally")
install.packages("PerformanceAnalytics")
install.packages("gmodels")
install.packages("RColorBrewer")
```
Sometimes this does not work, in which case go to tools>install packages>paste the following under CRAN: 
ggplot2,dplyr,dbplyr,foreign,ggpubr,xtable,mgcv,tree,validate, lubridate
Then load the libraries below
```{r}

library(ggplot2)
library(dplyr)
library(dbplyr)
library(foreign)
library(ggpubr)
library(xtable)
library(mgcv)
library(tree)
library(validate)
library(lubridate)
library(GGally)
library(PerformanceAnalytics)
library(gmodels)

```


# 1. Organise and clean the data

## 1.1 Subset the data into the specific dataset allocated
 
*A description of the data set provided, its contents and which subset you should select is documented in the assessment brief at ???.pdf*  
*Use R code to correctly select the subset of data allocated. (5 marks)*  

```{r}
# Code to select the student specific team combination.  
# Feel free to change it including the data frame name.
# 
# my.2teams.df <- subset(analysis.data, teamID.x=="???" | teamID.x=="???")


BOSLAA = subset(CS5801.data, teamID.x == "BOS" | teamID.x=="LAA")
BOS = subset(CS5801.data, teamID.x == "BOS")
LAA = subset(CS5801.data, teamID.x == "LAA")
#I am also renaming the dataframe to "League" because I have a personal preference not to have punctuation in the names of my objects.
League = CS5801.data


```
Only the two teams need to be separated at this stage, but I also decivded to have subsets including only one team each. The reason is that later on, I will probably want to look at each on of them separately. This is because depending on the task at hand, different sized datasets will be more useful. **spoiler alert**, it is useful at the end of part 1.


## 1.2 Data quality analysis
 
*Provide a description of a comprehensive plan to assess the quality of the data, and document your findings.  Include all variables/columns (5 marks) from the data set and provide a full implementation (5 marks).  NB even if no data quality issues are identified you should still check and report.*

In terms of quality, the data appears to be rather robust. We have more or less the same number of players per team (38 and 42).  When plotting them, one sees that they look quite similar too. 
Simply plotting a line shows that there are clear and observable trends.A summary shows anomalies very quickly. 
One anomaly is a negative number of games, runs, or length of career, for example. These are physical impossibilities. Another anomaly is that some things have a median of zero, or other extreme distributions.It would be interesting to see why.
Some of these anomalies do not appear to be in my teams, but let's take a look.

```{r}
summary(League)
plot(League, panel = panel.smooth)

summary(BOSLAA)
plot(BOSLAA, panel = panel.smooth)

```

It contains 15 variables. They are as follows:
1 - playerID
2 - teamID.x
3 - R
4 - G
5 - H
6 - AB
7 - RBI
8 - weight
9 - height
10 - salary
11 - birthDate
12 - career.legth
13 - bats
14 - age
15 - hit.ind



structure() and glimpse() are good here. This is because they tell me the size and appearance on the values as well as their type. I have mostly numeric (int and dbl) data, but also factors, characters, and a date
```{r}


structure(BOSLAA)
glimpse(BOSLAA)
```

I will also check for missing values. Luckily there are none here. Because the test prints a maximum number of values, splitting the subset into two (divided by team) allowed for all 80 rows to be visible. As can be seen, there is not a single missing value. Producing a table also shows that all 1200 values have been filled in.


```{r}

table(is.na(BOSLAA))
(is.na(BOS))
(is.na(LAA))


```
Returning rows with missing data also produces nothing. There are several ways to check for missing values. Here is another:
```{r}

BOSLAA[!complete.cases(BOSLAA),]

```
## 1.3 Data cleaning  
 
*Explain any data quality issues found in 1.2 (5 marks), justify and document the responses made (if any) (5 marks).*

The dataset has been renamed to League, because it carries more meaning. not only this, but later on, when comparing the two teams to the league as a whole, there will be fewer formatting issues. The reason is that punctuation in the wrong places can impact things, and having punctuation in the middle of a name can also be confusing to the reader as it could look like it has a purpose beyond beying a character in a name. It is also a personal stylistic choice.
I will rename some collumns as well, so that I can better understand what I am loking at at a glance, but also for the reasons above. While renaming, I can also separate it by team. While it looks like a lot of work, this will save a lot of time later on. It also means that my definitions are, for the large part, in one place.

The 15 variables will be renamed as follows:
platerID
teamID.x - Team
R - Runs
G - Games
H - Hits
AB - AtBat
RBI - RunsBatted
weight - Weight
height - Height
salary - Salary
birthDate - Birth
career.length - Career
bats - Bats
age - Age
hit.ind - SeasonHit. I do not know what hit.ind means, but it represents whether a player had a hit that season.It has a Bernoulli disctribution, which has a binary distribution. SeasonHit 1 or 0 basically turns it into a yes/no question which makes more intuitive sense to me.
The rest will simply be capitalised for consistent formatting
```{r}



League %>% 
  rename(
    Team = teamID.x,
    Runs = R,
    Games = G,
    Hits = H,
    AtBat = AB,
    RunsBatted = RBI,
    Weight = weight,
    Height = height,
    Salary = salary,
    Birth = birthDate,
    Career = career.length,
    Bats = bats,
    Age = age,
    SeasonHit = hit.ind
    )
BOSLAA %>% 
  rename(
    Team = teamID.x,
    Runs = R,
    Games = G,
    Hits = H,
    AtBat = AB,
    RunsBatted = RBI,
    Weight = weight,
    Height = height,
    Salary = salary,
    Birth = birthDate,
    Career = career.length,
    Bats = bats,
    Age = age,
    SeasonHit = hit.ind
    )
BOS %>% 
  rename(
    Team = teamID.x,
    Runs = R,
    Games = G,
    Hits = H,
    AtBat = AB,
    RunsBatted = RBI,
    Weight = weight,
    Height = height,
    Salary = salary,
    Birth = birthDate,
    Career = career.length,
    Bats = bats,
    Age = age,
    SeasonHit = hit.ind
    )
LAA %>% 
  rename(
    Team = teamID.x,
    Runs = R,
    Games = G,
    Hits = H,
    AtBat = AB,
    RunsBatted = RBI,
    Weight = weight,
    Height = height,
    Salary = salary,
    Birth = birthDate,
    Career = career.length,
    Bats = bats,
    Age = age,
    SeasonHit = hit.ind
    )



```
Here I am making subests of everything so that I do not have to later. Having this in one place means that I need to look in fewer places when making changes.
```{r}
League = CS5801.data

#1
LeagueGames = League$G
BOSLAAGames = BOSLAA$G
BOSGames = BOS$G
LAAGames = LAA$G
#2
LeagueRuns = League$R
BOSLAARuns = BOSLAA$R
BOSRuns = BOS$R
LAARuns = LAA$R
#3
LeagueHits = League$H
BOSLAAHits = BOSLAA$H
BOSHits = BOS$H
LAAHits = LAA$H
#4
LeagueAtBat = League$AB
BOSLAAAtBat = BOSLAA$AB
BOSAtBat = BOS$AB
LAAAtBat = LAA$AB
#5
LeagueRunsBatted = League$RBI
BOSLAARunsBatted = BOSLAA$RBI
BOSRunsBatted = BOS$RBI
LAARunsBatted = LAA$RBI
#6
LeagueWeight = League$weight
BOSLAAWeight = BOSLAA$weight
BOSWeight = BOS$weight
LAAWeight = LAA$weight
#7
LeagueHeight = League$height
BOSLAAHeight = BOSLAA$height
BOSHeight = BOS$height
LAAHeight = LAA$height
#8
LeagueSalary = League$salary
BOSLAASalary = BOSLAA$salary
BOSSalary = BOS$salary
LAASalary = LAA$salary
#9
LeagueBirth = League$birthDate
BOSLAABirth = BOSLAA$birthDate
BOSBirth = BOS$birthDate
LAABirth = LAA$birthDate
#10
LeagueCareer = League$career.length
BOSLAACareer = BOSLAA$career.length
BOSCareer = BOS$career.length
LAACareer = LAA$career.length
#11
LeagueBats = League$bats
BOSLAABats = BOSLAA$bats
BOSBats = BOS$bats
LAABats = LAA$bats
#12
LeagueAge = League$age
BOSLAAAge = BOSLAA$age
BOSAge = BOS$age
LAAAge = LAA$age
#13
LeagueSeasonHit = League$hit.ind
BOSLAASeasonHit = BOSLAA$hit.ind
BOSSeasonHit = BOS$hit.ind
LAASeasonHit = LAA$hit.ind
#14
LeagueplayerID = League$playerID
BOSLAAplayerID = BOSLAA$playerID
BOSplayerID = BOS$playerID
LAAplayerID = LAA$playerID
#15
#The team names are not here, because only two will be used, and they are present in all of the subsets above.
```
Before taking any further steps, it is clear that some values make no sense. For example, BOS has a player with a negative RBI. It is not possible to run a negative number of runs. check_that() is very useful as it allows me to create rules and check them. Rules such as valid answers, or logical answers. This is first done for the league as a whole, going down in number. 

I might have to turn these subsets back into dataframes, but at least I understand what the figures mean. This will be important when analysing everything later.
It might seem like a good idea to check for duplicate players here. This is not clever. The reason is that the same player can play for multiple teams.

```{r}
## Data quality checking for the whole league

summary(League)

LeagueCheck <- check_that(League, 
                          LeagueRuns>= 0,
                          LeagueGames>= 0,
                          LeagueHits >= 0,
                          LeagueAtBat >= 0,
                          LeagueRunsBatted >= 0,
                          LeagueWeight > 0,
                          LeagueHeight > 0,
                          LeagueSalary > 0,
                          LeagueBirth > 0,
                          LeagueCareer >= 0,
                          LeagueAge > 0,  
                          LeagueBats == "R" | LeagueBats == "L" | LeagueBats == "B",
                          LeagueSeasonHit == "1" | LeagueSeasonHit == "0")
#Duplicate players are not looked at here for a reason     
# Produce a bar chart of the quality rule failures
barplot(LeagueCheck, main = "League Verification")


```
I see very quickly that there are four areas where the rules were broken. This is because they are negative, and in this case any negative values are a physical impossiblity. These four are:
LeagueCareer
LeagueRunsBatted
LeagueGames
LeagueRuns
These foud have at least one entry which is negative.

Now I need to see whether they are present in either team:

```{r}

summary(BOSLAA)

BOSLAACheck <- check_that(BOSLAA, 
                          BOSLAACareer>= 0,
                          BOSLAARunsBatted>= 0,
                          BOSLAAGames >= 0,
                          BOSLAARuns >= 0)
             
# In order to roduce a bar chart of the quality rule failures. I need to turn it into a dataframe   

barplot(BOSLAACheck, main="Boston and LA Verification")

```
Here I see that there is one negative RBI. This would appear to be the only problematic value at this stage.

```{r}
LAACheck <- check_that(LAA, 
                          LAARunsBatted>= 0)

barplot(LAACheck, main = "LA Verification")

BOSCheck <- check_that(BOS, 

                          BOSRunsBatted>= round(0))
barplot(BOSCheck, main = "Boston Verification")
BOSCheckDF =  as.data.frame(BOSCheck)
#let's look at the problem player
ProblemPlayer <- subset(BOSCheckDF, value ==FALSE)
ProblemPlayer
#I have located the player but not identified it

```
The RBI problem is in Boston. This error could have various causes. I have no way of telling the true value at this stage, however. the reason why, is that the relevant dataframe looks are 3 variables, and none of these match the 15 from BOSLAA Name, value and expression have no analogues in BOSLAA. I can search for the player manually, and their name is grokkel01. They d not have any other unusual values.

```{r}
str(BOSCheckDF) 
```

Now to identify the duplicate players.
```{r}

BOSLAADouble = BOSLAA$playerID[duplicated(BOSLAA$playerID)]
BOSLAADouble

BOSDoubles = duplicated(BOSplayerID) 
table(BOSDoubles)


LAADoubles = duplicated(LAAplayerID) 
table(LAADoubles)


```

Each team seems to have a duplicate player. BOS has pedrodu01, and LAA has richaga01. There are three duplicate players in total, however. 

Now to remove the duplicated player from each team. this will be done on each team individually, this is because one player could have played on both teams in the same year.
```{r}
#I need to be careful to only remove duplicates in terms of name, and nothing else.

BOSClean <- BOS[!duplicated(BOS$playerID), ]

LAAClean <- LAA[!duplicated(LAA$playerID), ]
BOSClean
LAAClean
#here I remove each duplicate from each team
BOSCleanDoubles = duplicated(BOSClean$playerID) 
table(BOSCleanDoubles)

LAACleanDoubles = duplicated(LAAClean$playerID) 
table(LAACleanDoubles)
# here I bind the two clean teams together
BOSLAAClean = rbind(BOSClean, LAAClean)
BOSLAAClean
```
I can also check for duplicate players here in the clean dataset. It looks like victosh01 played for both teams. his performance was actually quite similar across both teams. It is not identical. I previously said that it was possible for one player to playa cross two teams. This is an example.

```{r}
BOSLAACleanDoubles = duplicated(BOSLAAClean$playerID) 
table(BOSLAACleanDoubles)
BOSLAACleanDoubles
BOSLAACleanDouble = BOSLAAClean$playerID[duplicated(BOSLAAClean$playerID)]
BOSLAACleanDouble
BOSLAAClean
```

I can now comfortably look at the data, knowing that I need to watch out for this single negative RBI value. Hopefully the reason for its being will be revealed.
I need a clean version of everything now.

```{r}
#1
BOSLAACleanGames = BOSLAAClean$G
BOSCleanGames = BOSClean$G
LAACleanGames = LAAClean$G
#2
BOSLAACleanRuns = BOSLAAClean$R
BOSCleanRuns = BOSClean$R
LAACleanRuns = LAAClean$R
#3
BOSLAACleanHits = BOSLAAClean$H
BOSCleanHits = BOSClean$H
LAACleanHits = LAAClean$H
#4
BOSLAACleanAtBat = BOSLAAClean$AB
BOSCleanAtBat = BOSClean$AB
LAACleanAtBat = LAAClean$AB
#5
BOSLAACleanRunsBatted = BOSLAAClean$RBI
BOSCleanRunsBatted = BOSClean$RBI
LAACleanRunsBatted = LAAClean$RBI
#6
BOSLAACleanWeight = BOSLAAClean$weight
BOSCleanWeight = BOSClean$weight
LAACleanWeight = LAAClean$weight
#7
BOSLAACleanHeight = BOSLAAClean$height
BOSCleanHeight = BOSClean$height
LAACleanHeight = LAAClean$height
#8
BOSLAACleanSalary = BOSLAAClean$salary
BOSCleanSalary = BOSClean$salary
LAACleanSalary = LAAClean$salary
#9
BOSLAACleanBirth = BOSLAAClean$birthDate
BOSCleanBirth = BOSClean$birthDate
LAACleanBirth = LAAClean$birthDate
#10
BOSLAACleanCareer = BOSLAAClean$career.length
BOSCleanCareer = BOSClean$career.length
LAACleanCareer = LAAClean$career.length
#11
BOSLAACleanBats = BOSLAAClean$bats
BOSCleanBats = BOSClean$bats
LAACleanBats = LAAClean$bats
#12
BOSLAACleanAge = BOSLAAClean$age
BOSCleanAge = BOSClean$age
LAACleanAge = LAAClean$age
#13
BOSLAACleanSeasonHit = BOSLAAClean$hit.ind
BOSSCleaneasonHit = BOSClean$hit.ind
LAACleanSeasonHit = LAAClean$hit.ind
#14
BOSLAACleanplayerID = BOSLAAClean$playerID
BOSCleanplayerID = BOSClean$playerID
LAACleanplayerID = LAAClean$playerID
```



        
# 2. Exploratory Data Analysis (EDA)



## 2.1 EDA plan

*Outline a suitable plan to explore, describe and visualise your data. (5 marks)*  

i will undergo a process including two cyclces which I will switch between at each stage of the data analysis. Most people will use a similar process. This one is inspired by Peng and Matsui's method in their book, **the Art of Data Science**. They call them epicycles.
The first cycle will be to look at the data, develop assumptions, and then see whether they match.
The second cycle will be ask the questions relevant to these assumptions, see what models might be appropriate, build them, interpret the results, and then report them.

The first epicycle was more or less done in the data cleaning stage. Now I need to build some models.  I will start with some histograms and bar graphs for all data, as well as box plots. Not only this, but means and medians where applicable. I will try to see if any are skewed, and use a Poisson method to measure this.
I will then look at the new data again

I will do a summary and histogram of each variable aside from birthdate, and batting hand. This is because they are dates and factors which are not suitable for this format.
Because this is a relatively small sample size, rather than focus on standard deviation, I will calculate the t as well. This is also because I do not know what to expect in terms of the standard deviation.



## 2.2 EDA and summary of results  

*Undertake and summarise the findings of your data exploration, particularly with respect to the research questions.  Use appropriate summary statistics (uni- and multi-variate) and visualisations. (10 marks)*

I will do a histogram of each variable with a density line. This will allow me to identify any clear trends or distributions, and therefore infer any dependencies.
Batting hand will not be graphed this way because it is a factor. Birthdate will not either.

```{r}
##variablew platerID
##teamID.x
##R - Runs
##G - Games
##H - Hits
##AB - AtBat
##RBI - RunsBatted
##weight 
##height
##salary
##birthDate - Birth
##career.length - Career
##bats
##age
##hit.ind - SeasonHit

summary(BOSLAAClean)

hist(BOSLAACleanRuns, prob= T)
lines(density(BOSLAACleanRuns), col = "red")
hist(BOSLAACleanGames, prob= T)
lines(density(BOSLAACleanGames), col = "red")
hist(BOSLAACleanHits, prob= T)
lines(density(BOSLAACleanHits), col = "red")
hist(BOSLAACleanAtBat, prob= T)
lines(density(BOSLAACleanAtBat), col = "red")
hist(BOSLAACleanRunsBatted, prob= T)
lines(density(BOSLAACleanRunsBatted), col = "red")
hist(BOSLAACleanWeight, prob= T)
lines(density(BOSLAACleanWeight), col = "red")
hist(BOSLAACleanHeight, prob= T)
lines(density(BOSLAACleanHeight), col = "red")
hist(BOSLAACleanSalary, prob= T)
lines(density(BOSLAACleanSalary), col = "red")
hist(BOSLAACleanCareer, prob= T)
lines(density(BOSLAACleanCareer), col = "red")
hist(BOSLAACleanAge, prob= T)
lines(density(BOSLAACleanAge), col = "red")
hist(BOSLAACleanSeasonHit, prob= T)
lines(density(BOSLAACleanSeasonHit), col = "red")


```

This is quite interesting, because we have a nice assortment of different kinds of disctributions. 


There are many things that instantly stand out which I will be able to test soon. 
Right now, I will look at the distributions.

Height, weight, and age appear to have binomial distribution.

Runs, hits, bats, games, salary, career and at bats all seem to have poisson distribution. This is because evrybody starts at zero.

Season hit has a bernouilli distribution.

Now is time to test these.

```{r}

shapiro.test(BOSLAACleanWeight)
shapiro.test(BOSLAACleanHeight)
shapiro.test(BOSLAACleanAge)



```
Age, weight, and height are normal, as their p-value is over 0.05. This means that there is indeel a normal disribution.
The others need to be looked at.

I can also measure the skewness of Runs, hits, runs batted, games, salary, career. They are indeed skewed.

```{r}

RunSkew = skewness(BOSLAACleanRuns)
RunSkew
HitSkew = skewness(BOSLAACleanHits)
HitSkew
RunsBattedSkew = skewness(BOSLAACleanRunsBatted)
RunsBattedSkew
AtBatSkew = skewness(BOSLAACleanAtBat)
AtBatSkew
GamesSkew = skewness(BOSLAACleanGames)         
GamesSkew
SalarySkew = skewness(BOSLAACleanSalary)
SalarySkew
CareerSkew = skewness(BOSLAACleanCareer)         
CareerSkew
```
Skewness is greater than one for all of these measures as it is greater than 1 for all values. The most extreeme are career length at the most skewed, and the games playes at the least skewed. It is pretty clear why career length  would be more skewed than games played in a season. The reason is that everybody is able to achieve the maximum number of games in a given season in theory, while career length is can have more extreme values. We also leran here that baseball players probably have a relativaly long career lifespan, given that tehy are sportspeople. 
## 2.3 Additional insights and issues

*Highlight potential further issues or insights uncovered in 2.2.  This might include follow up to findings from your initial EDA.  We accept that the boundary between 2.2 and 2.3 is somewhat arbitrary so use your judgement and maximise good structure and readability. (5 marks)*

Seeing as the median is zero in a few graphs, yielding extremely skewed distributions, to remove the zero values would allow me to learn a lot more about the rest. Let's take a look at Runs and RunsBatted.
In fact, it would also be helpful to do the same to Hits and AtBat
What you have here is what Nassim Taleb might call a fat-tailed distribution in these cases. This means that given that everybody  starts at zero, but the upper limit is theoretically unlimited. What will happen in this case is precicely what is found here.

```{r}
## create new sets removing values =0
BOSLAACleanRunsDone = BOSLAACleanRuns[BOSLAACleanRuns != "0"]
hist(BOSLAACleanRunsDone, prob = T)
lines(density(BOSLAACleanRunsDone), col = "red")
summary(BOSLAACleanRunsDone)

BOSLAACleanHitsDone = BOSLAACleanHits[BOSLAACleanHits != "0"]
hist(BOSLAACleanHitsDone, prob = T)
lines(density(BOSLAACleanHitsDone), col = "red")
summary(BOSLAACleanHitsDone)

BOSLAACleanAtBatDone = BOSLAACleanAtBat[BOSLAACleanAtBat != "0"]
hist(BOSLAACleanAtBatDone, prob = T)
lines(density(BOSLAACleanAtBatDone), col = "red")
summary(BOSLAACleanAtBatDone)

BOSLAACleanRunsBattedDone = BOSLAACleanRunsBatted[BOSLAACleanRunsBatted !="0"]

##This is a very good moment to remove that -4
hist(BOSLAACleanRunsBattedDone, prob = T)
lines(density(BOSLAACleanRunsBattedDone), col = "red")
summary(BOSLAACleanRunsBattedDone)

##Another way to do this is to select sums greater than or lower than a certain amount. I will probably to this to analyse those who did not do any runs hits, at bats, or bats.
RunsDoneSkew =  skewness(BOSLAACleanRunsDone)
RunsDoneSkew
HitsDoneSkew = skewness(BOSLAACleanHitsDone)
HitsDoneSkew
AtBatDoneSkew = skewness(BOSLAACleanAtBatDone) 
AtBatDoneSkew
RunsBattedDoneSkew = skewness(BOSLAACleanRunsBattedDone)
RunsBattedDoneSkew
##
```
We see that the distribution is still skewed in all three aside from the runs which looks more normal. BOSLAACleanRunsBattedDone  might seem like something interesting to take away. Looking carefully, there is a negative value. This is a good time to make it a little bit cleaner. I cheated by turning the negative value positive by using the abs() function. I can do this with confidence because the other figures present are not particularly unusual. Skewness here, falls from above one to below one, taking from highly to moderately skewed.
Another thing to now is that skewness for runsdone, hitsdone, and atbats done hasfalled from above 1.5 to below 1, which is from highly skewed to moderately skewed.
This is the guidance I used: https://help.gooddata.com/doc/en/reporting-and-dashboards/maql-analytical-query-language/maql-expression-reference/aggregation-functions/statistical-functions/predictive-statistical-use-cases/normality-testing-skewness-and-kurtosis

```{r}
BOSLAACleanerRunsBattedDone <- abs(BOSLAACleanRunsBattedDone)

BOSLAACleanerRunsBattedDone

hist(BOSLAACleanerRunsBattedDone, prob = T)
lines(density(BOSLAACleanerRunsBattedDone), col = "red")
summary(BOSLAACleanerRunsBattedDone)
CleanerRunsBattedDoneSkew = skewness(BOSLAACleanerRunsBattedDone)
CleanerRunsBattedDoneSkew
```

The cleaner version shows that this is clearly not a normal distribution. Although it might appear to be less skewed, it is more skewed thanRuns, Bats, Hits, or AtBats.
Thi s is interesting because runsbatted is less impacted by the numbers of bats, hits, or at bats when removing zero values. 
Understanding why makes sense takes a little bit of thought. The reason why players ho do not have batted runs is less impacted by those with zero is that runs batted are almost inevitably lower, or maximally equal to the number of runs, hits, or bats. Because the values are less extreme, there is less skewness, and therefore less room to move.


This is very interesting. Although career length had amonth the highest median values in terms of the skewed (by virtue of career length's impossiblitliy of being zero), career length is the most skewed. This is  further accentuated when removing the zero values from runs, games, hit, and at bats, all of which lower skewness.


# 3. Modelling

## 3.1 Build a model for player salary

*Given the research question (i.e., the target attribute of salary) outline an analysis plan that incorporates/references any findings from the data cleaning (1.3) and EDA (2.2)  (5 marks). Use R to build a suitable model (10 marks).*  
*NB Submissions where suitable models do not have good fit due to the nature of the data will not be penalised.*  

The first thing that I can do is look at all of the collelants.This is quite good, because I can see in one shot what appears to affect what. Some of this might be misleading. For example, height and weight seem to have little to do with salary. This is to do with the fact that height and weight have a normal distribution. There could be an ideal height and weight, and this is not indicated here. The strongest relationship seems to do with age and career length. Then, whether they scored a hit. Height, weight, and runs batted come next. Finally, runs, games, hits, and whether they batted seem to bear nearly no relation with salary.


```{r}
ggcorr(BOSLAAClean, nbreaks = 20)
#although the number of breaks might seen high, it allows for the accuracy of a continuous gradient while being able to know what value it corresponds to.
#Both have been included for comparison
ggcorr(BOSLAAClean)

```
I wanted to add a colourblind version of this for accessibility, but because there are non numeric arguments in the dataset, it is not possible to make a colourblind version without different methods.
```{r}
display.brewer.all(colorblindFriendly = TRUE)
ggcorr(BOSLAAClean + scale_color_brewer(palette = "Dark2"))
```

I want to see the correlates between each variable and player salary. Then, within that, see wether any of those items move win or against one another. Such as career length and age. or runs and bats. This is to see whether I am correct to assume that these elements might be connected.
ANOVA and ANCOVA will be useful for finding a lot of information in one go.
Another thing is that I will go one by one and remove the weakest correlants.



For example, a box plot will  be useful for the season hit statistic in particular.


What I learn here is that batting hand and salary seem to be related somehow.
For example, a box plot will  be useful for the season hit, and batting hand statistics in particular.



```{r}


ggplot(BOSLAAClean, aes(x=BOSLAACleanBats, y=BOSLAACleanSalary)) + geom_boxplot()
ggplot(LAAClean, aes(x=LAACleanBats, y=LAACleanSalary)) + geom_boxplot()
ggplot(BOSClean, aes(x=BOSCleanBats, y=BOSCleanSalary)) + geom_boxplot()
```
Handedness does not seem to have a huge impact, while left-handed player seem to be at a slight disadvantage. While ambidextrous seem to earn more than right handed people and even more than left handed people, the latter two have a very large number of outliers.

What about career lenght, and age?
What about birthmonth and age?
I will also look at birth month and salary.The reason for this is that there are many studies which show that the birth month of a person can have a high impact on their sports performance. This is due to how the school year functions. I will also check the year, while I'm at it.
Thi s is only one study showing the birth month theory: http://www.ncaa.org/about/resources/research/birthday-effect-college-athletics
It suggests that August through November are the best period, followed by December to March, and finally April to July.
This is another such study: Athletic Performance and Birth Month: Is the Relative Age Effect More than just Selection Bias?
, DOI: 10.1055/s-0034-1368725
For this to work, I need to extract the month from the birth date, and then turn it into a factor so that it they can be separate like in the batting box plot above. The same will be done with season hit which has only two possible answers, 0 and 1.

```{r}
BOSLAACleanMonth = month(BOSLAACleanBirth)

BOSLAACleanMonthFactor = factor(BOSLAACleanMonth)
levels(BOSLAACleanMonthFactor)
ggplot(BOSLAAClean, aes(x=BOSLAACleanMonthFactor, y=BOSLAACleanSalary)) + geom_boxplot() + geom_jitter(color = "red")

BOSLAACleanYear = year(BOSLAACleanBirth)

BOSLAACleanYearFactor = factor(BOSLAACleanYear)
levels(BOSLAACleanYearFactor)
ggplot(BOSLAAClean, aes(x=BOSLAACleanYearFactor, y=BOSLAACleanSalary, prob = T,)) + geom_boxplot() +
geom_jitter(color = "red")

```
The month when one was born does not seem to have a huge impact, although being born either at the bening or end of the year seems to result in higher salaries than in the middle. The advantage seems to lie with being born in the second half of the year.
This seems to corroborate with the findings from the NCAA which concluded that the end of the year was strongest, followed by the start, and finally the middle being weakest.
The year of birth, however, shows a clear trend that the longer one plays, the higher the salary.

How much does whether one scored a hit impact salary?
To do this, I am using a boxplot and this involves turning the numbers into factors. 0 being no, and 1 being yes.
```{r}



BOSLAACleanSeasonHitFactor = factor(BOSLAACleanSeasonHit)
levels(BOSLAACleanSeasonHitFactor)
ggplot(BOSLAAClean, aes(x=BOSLAACleanSeasonHitFactor, y=BOSLAACleanSalary)) + geom_boxplot()

```
This is not very surprising. It would seems as though pay is related to performance. Some people still earn quite a lot without scoring a single hit. It could be tat they are catchers or something else which does not involve batting. The reality is that I do not know.

Obvious things to test whether they correlate are career length and salary, especially since we saw that the birth year seems linked. We need to see whether it is only a function of length, or seniority.
```{r}
##testing correlations is important here
cor(BOSLAACleanAge, BOSLAACleanSalary, method = c("pearson", "kendall", "spearman"))
cor.test(BOSLAACleanAge, BOSLAACleanSalary, method=c("pearson", "kendall", "spearman"))
cor(BOSLAACleanCareer, BOSLAACleanSalary, method = c("pearson", "kendall", "spearman"))
cor.test(BOSLAACleanCareer, BOSLAACleanSalary, method=c("pearson", "kendall", "spearman"))
cor(BOSLAACleanSeasonHit, BOSLAACleanSalary, method = c("pearson", "kendall", "spearman"))
cor.test(BOSLAACleanSeasonHit, BOSLAACleanSalary, method=c("pearson", "kendall", "spearman"))
cor(BOSLAACleanMonth, BOSLAACleanSalary, method = c("pearson", "kendall", "spearman"))
cor.test(BOSLAACleanMonth, BOSLAACleanSalary, method=c("pearson", "kendall", "spearman"))

```


```{r}
#the original plan was to only graph what was done, but it will not happen.

HitsLM = (lm(formula = BOSLAACleanHits ~ BOSLAACleanSalary))
plot(HitsLM)
GamesLM = (lm(formula = BOSLAACleanGames ~ BOSLAACleanSalary))
plot(GamesLM)
RunsLM = (lm(formula = BOSLAACleanRuns ~ BOSLAACleanSalary))
plot(RunsLM)
RunsBattedLM = (lm(formula = BOSLAACleanRunsBatted ~ BOSLAACleanSalary))
plot(RunsBattedLM)


```
Step funcions are also useful here
```{r}
step(GamesLM)
step(RunsLM)
step(RunsBattedLM)
```


```{r}

summary(GamesLM)
summary(RunsLM)
summary(RunsBattedLM)
```

The next thing that I want to look at is height and weight. Is there an ideal height or weight?

```{r}
HeightLM = lm(formula = BOSLAACleanHeight ~ BOSLAACleanSalary)
plot(HeightLM)
WeightLM = lm(formula = BOSLAACleanWeight ~ BOSLAACleanSalary)
plot(WeightLM)
#create model comparing heigth and weright - scatter plot?
```
Looking here, it seems as though there is an ideal height and weight as on th eQ-Q plot, the points are closer to the line towards the middle.
```{r}
step(HeightLM)
step(WeightLM)
```
```{r}
summary(HeightLM)
summary(WeightLM)
```



```{r}
SalaryLM = (lm(formula = BOSLAACleanSalary ~ BOSLAACleanSalary + BOSLAACleanHits + BOSLAACleanGames + BOSLAACleanRuns +BOSLAACleanHeight + BOSLAACleanWeight))
plot(SalaryLM)
step(SalaryLM)
```



## 3.2 Critique model using relevant diagnostics

*Offer an interpretation of the model characteristics, goodness of fit and graphical diagnostics (5 marks) for the model built in 3.1. Explain any potential weaknesses (5 marks).*
What we find here is pretty amazing. Career length and age have very similar correlations. Older players, it would seem, seem to have played longer. This makes sense as people usualy start sports at around the same age.
The birth month has a much weaker correlation.
We can learn several things. One is that the longer a person plays, the more they will make, and it does not matter what year they start in. This is not about age of starting, but the calendar year when they started. It could be possible that players from one generation earn more or less than the other due to perhaps historical or cultural factors. If there are, they probably are not very significant. More important is the amount of time played, than time of tear born.

This is very interesting. 

Now that we see that the longer someone has played indicates a higher salary, what about performance during the season itself?
It might seem obvious that performance and salary would correlate, but not necessarily. For instance, studies have shown that athletes' performance can dip after transferring from one team to another. This transfer usually follows an offer of a higher salary. There was another stuyd in Israel as noted by Judea Pearl in the Book of Why about the effect of praise on soldier performances. It seemed to have an inverse correlation, but this was due to some form of mean regression and the praise or criticism had little or no impact.

Because some players will never bat or run because players on a baseball team sometimes have specialisations, I woud like to remove all players with zero values, but will not. This is a complicated decision because some of the players with zero will be unnecessary, others will be necessary. The proportion of these is unknown without knowing more about each individual player. It is important to compare like-for-like. It would not be correct to measure salary on batting performance on someone who is not on the team to bat, for example. They are being paid for something else otherwise undefined. The problem is that none of the roles have been defined. Due to this, I will not remove any players.
Another complication is that removing players with zero values results in datasets of different sizes, which makes comparison more complicated.
The main weakness is that I have not ben able to divide the populations adequately. I need to be able to know which player played in which position, and divide them accordingly. The reason is that different players perform differently based on which position that they are in.






## 3.3 Suggest improvements to your model

*Based on the findings in 3.2 articulates possible alternative approaches to address them (5 marks).*
The easiest and most obvious way to improve the model would be to have even more data to compare to. This will allow one to track chenges over time. Player salary could behave differently according to the player profile over time. Some players would have high salaries over a short period, while others could have high salaries over a long period.
Luckily, I do have more data. I can compare it with the league as a whole. Not only this, but I could also get a hold of data over time.

Because some players will never bat or run because players on a baseball team sometimes have specialisations, I woud like to remove all players with zero values, but will not. This is a complicated decision because some of the players with zero will be unnecessary, others will be necessary. The proportion of these is unknown without knowing more about each individual player. It is important to compare like-for-like. It would not be correct to measure salary on batting performance on someone who is not on the team to bat, for example. They are being paid for something else otherwise undefined. The problem is that none of the roles have been defined. Due to this, I will not remove any players.
Another complication is that removing players with zero values results in datasets of different sizes, which makes comparison more complicated.


I need training data in order to compare my findings.
I also need to be able to creat upper and lower confidence limits.

I want to dig little bit deeper and look at the variance as well. This is a way to test my salary model against something larger.


# 4. Extension work

## 4.1 Model the likelihood of a player having scored a Hit (using the hit.ind variable provided).

*Given this second research question (i.e., involving the binary target attribute) provide a plan of analysis based on relevant EDA for this attribute (10 marks). The model is described, explained and critiqued (10 marks).*
*NB Submissions where suitable models do not have good fit due to the nature of the data will not be penalised.* 

This will be done on four levels. The whole set, two teams, and then each of the two teams. This iwll be very similar to how the model for player salary was developed.

Because hit.ind/seasonhit is a binary variable, logistic regression is the best model according to the lectures.
##use logistic regression here somehow

```{r}
table(BOSLAACleanSeasonHit)
mean(BOSLAACleanSeasonHit)

```

```{r}
CrossTable(BOSLAACleanSeasonHit, BOSLAACleanBats, digits=2, prop.r = TRUE, prop.c = TRUE,  prop.chisq = FALSE, 
           chisq = TRUE, fisher = TRUE)
```
##crosstable shows more information - i don't knwo how to use it , however.
```{r}
HitsLM = lm(formula = BOSLAACleanSeasonHit ~ BOSLAACleanSalary + BOSLAACleanHits + BOSLAACleanGames + BOSLAACleanRuns +BOSLAACleanHeight + BOSLAACleanWeight)
plot(HitsLM)
step(HitsLM)
```


Using 
```{r}
LeagueHitsGLM<-glm(hit.ind~G+R+H+AB+RBI+weight+height+career.length+bats+salary ,data = League, family = "binomial")
LeagueHitsGLM
###this gives really low numbers for some reason
```
```{r}
step(LeagueHitsGLM)

```

```{r}
BOSLAACleanHitsGLM<-glm(hit.ind~G+R+H+AB+RBI+weight+height+career.length+bats+salary ,data = BOSLAA, family = "binomial")
BOSLAACleanHitsGLM
###this gives really low numbers for some reason
```

```{r}
step(BOSLAACleanHitsGLM)
summary.lm(BOSLAACleanHitsGLM)
```
after this, I can arrive at a minimal model

use predict function afterwards

The source for using this is lectures 6 and 7


# References  



*Add any references here. NB You can either do this manually or automatically with a `.bib` file (which then must be submitted along with your `.Rmd` file).  See the RMarkdown [documentation](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html) for guidance.*    
