---
title: "CS5801 Coursework Template Proforma"
author: "student-id"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
version: 1.0
---

# 0. Instructions 

*1. Remove the (italicised) guidance text but keep the section headings.*  
*2. Add as many chunks of R code as required.*  
*3. Add descriptions of your analysis plans and explanations of your code and findings.  Please be detailed and where you have made choices explain the rationale for them.*  
*4. Write your report using RMarkdown.  For guidance see a [helpful blog](https://www.dataquest.io/blog/r-markdown-guide-cheatsheet/#tve-jump-17333da0719) or use the R Markdown cheatsheet which can be accessed from within RStudio by selecting `Help > Cheatsheets > R Markdown Cheat Sheet`.*  
*5. Your report should be clearly and professionally presented with appropriate use of cited external sources. (5 marks)*  
*6. It should also be easy to understand, with well-documented code following the principles of literate programming. (5 marks)*  

```{r}
# Add code here to load any required libraries with `library()`.  
# We suggest you use `install.package()` for any required packages externally to this document 
# since installation only need be done once.

install.packages("ggplot2")
install.packages("dplyr")
install.packages("dbplyr")
install.packages("foreign")
install.packages("ggpubr")
install.packages("xtable")
install.packages("mgcv")
install.packages("tree")
install.packages("validate")
install.packages("ggpubr")
```
Sometimes this does not work, in which case go to tools>install packages>paste the following under CRAN: ggplot2,dplyr,dbplyr,foreign,ggpubr,xtable,mgcv,tree,validate
Then load the libraries below
```{r}

library(ggplot2)
library(dplyr)
library(dbplyr)
library(foreign)
library(ggpubr)
library(xtable)
library(mgcv)
library(tree)
library(validate)
library("ggpubr")
```


# 1. Organise and clean the data

## 1.1 Subset the data into the specific dataset allocated
 
*A description of the data set provided, its contents and which subset you should select is documented in the assessment brief at ???.pdf*  
*Use R code to correctly select the subset of data allocated. (5 marks)*  

```{r}
# Code to select the student specific team combination.  
# Feel free to change it including the data frame name.
# 
# my.2teams.df <- subset(analysis.data, teamID.x=="???" | teamID.x=="???")
# Only the two teams need to be separated at this stage, but I also decivded to have subsets including only one team each. The reason is that 

BOSLAA = subset(CS5801.data, teamID.x == "BOS" | teamID.x=="LAA")
BOS = subset(CS5801.data, teamID.x == "BOS")
LAA = subset(CS5801.data, teamID.x == "LAA")

League = CS5801.data

summary(League)
summary(BOSLAA)
summary(BOS)
summary(LAA)



```


## 1.2 Data quality analysis
 
*Provide a description of a comprehensive plan to assess the quality of the data, and document your findings.  Include all variables/columns (5 marks) from the data set and provide a full implementation (5 marks).  NB even if no data quality issues are identified you should still check and report.*

In terms of quality, the data appears to be rather robust. We have more or less the same number of players per team.  When plotting them, one sees that they look quite similar too. When simply plotting shows that there are clear trends, and a summary shows a couple of anomalies. One is a negative number of games, runs, or length of career, for example. Another anomaly is that some things have a median of zero.


It contains 15 variables. They are as follows:
1 - platerID
2 - teamID.x
3 - R
4 - G
5 - H
6 - AB
7 - RBI
8 - weight
9 - height
10 - salary
11 - birthDate
12 - career.legth
13 - bats
14 - age
15 - hit.ind
structure() and glimpse() are good here
```{r}

structure(BOSLAA)
glimpse(BOSLAA)
```


```{r}
##here I am checking for missing values. Luckily there are none here. Because this prints a maximum number of values, splitting the subset allowed for all to be visible. As can be seen, there is not a single missing value.
is.na(BOS)
is.na(LAA)

```
I will also check this way:
is.na(filename)
finds missing values
str(surveyDF)
structure
glimpse
mean, median, quartile




## 1.3 Data cleaning  
 
*Explain any data quality issues found in 1.2 (5 marks), justify and document the responses made (if any) (5 marks).*

I will rename some collumns, so that I can better understand what I am loking at at a glance. This will help better understand the data as well. While renaming, I can also separate it by team. While it looks like a lot of work, this will save a lot of time later on.

It contains 15 variables. They will be renamed as follows:
platerID
teamID.x
R - Runs
G - Games
H - Hits
AB - AtBat
RBI - RunsBatted
weight 
height
salary
birthDate - Birth
career.length - Career
bats
age
hit.ind - SeasonHit

```{r}


#1
LeagueGames = League$G
BOSLAAGames = BOSLAA$G
BOSGames = BOS$G
LAAGames = LAA$G
#2
LeagueRuns = League$R
BOSLAARuns = BOSLAA$R
BOSRuns = BOS$R
LAARuns = LAA$R
#3
LeagueHits = League$H
BOSLAAHits = BOSLAA$H
BOSHits = BOS$H
LAAHits = LAA$H
#4
LeagueAtBat = League$AB
BOSLAAAtBat = BOSLAA$AB
BOSAtBat = BOS$AB
LAAAtBat = LAA$AB
#5
LeagueRunsBatted = League$RBI
BOSLAARunsBatted = BOSLAA$RBI
BOSRunsBatted = BOS$RBI
LAARunsBatted = LAA$RBI
#6
LeagueWeight = League$weight
BOSLAAWeight = BOSLAA$weight
BOSWeight = BOS$weight
LAAWeight = LAA$weight
#7
LeagueHeight = League$height
BOSLAAHeight = BOSLAA$height
BOSHeight = BOS$height
LAAHeight = LAA$height
#8
LeagueSalary = League$salary
BOSLAASalary = BOSLAA$salary
BOSSalary = BOS$salary
LAASalary = LAA$salary
#9
LeagueBirth = League$birthDate
BOSLAABirth = BOSLAA$birthDate
BOSBirth = BOS$birthDate
LAABirth = LAA$birthDate
#10
LeagueCareer = League$career.length
BOSLAACareer = BOSLAA$career.length
BOSCareer = BOS$career.length
LAACareer = LAA$career.length
#11
LeagueBats = League$bats
BOSLAABats = BOSLAA$bats
BOSBats = BOS$bats
LAABats = LAA$bats
#12
LeagueAge = League$age
BOSLAAAge = BOSLAA$age
BOSAge = BOS$age
LAAAge = LAA$age
#13
LeagueSeasonHit = League$hit.ind
BOSLAASeasonHit = BOSLAA$hit.ind
BOSSeasonHit = BOS$hit.ind
LAASeasonHit = LAA$hit.ind


#these names will late rbe added
#Runs = R
#Hits = H
#AtBat = AB
#RunsBatted = RBI
#LeagueHit= hist.ind

median(LeagueGames)
median(BOSLAAGames)
median(BOSGames)
median(LAAGames)

```
Before taking any further steps, it is clear that some values make no sense. For example, BOS has a player with a negative RBI. check_that() is very useful as it allows me to create rules and check them. Rules such as valid answers, or logical answers. This is first done for the league as a whole.  

I might have to turn these back into dataframes, but at least I understand what the figures mean. This will be important when doing aov and similar things

```{r}
## Data quality checking for the whole league

summary(League)

LeagueCheck <- check_that(League, 
                          LeagueRuns>= 0,
                          LeagueGames>= 0,
                          LeagueHits >= 0,
                          LeagueAtBat >= 0,
                          LeagueRunsBatted >= 0,
                          LeagueWeight > 0,
                          LeagueHeight > 0,
                          LeagueSalary > 0,
                          LeagueBirth > 0,
                          LeagueCareer >= 0,
                          LeagueAge > 0,  
                          LeagueBats == "R" | LeagueBats == "L" | LeagueBats == "B",
                          LeagueSeasonHit == "1" | LeagueSeasonHit == "0")
              
# Produce a bar chart of the quality rule failures
barplot(LeagueCheck)


```
I see very quickly that there are four values which do not make sense. This is because they are negative, and in this case any negative values are a physical impossiblity. These four are:
LeagueCareer
LeagueRunsBatted
LeagueGames
LeagueRuns
These foud have at least one entry which is negative.

Now I need to see whether they are present in either team:

```{r}

summary(BOSLAA)

BOSLAACheck <- check_that(BOSLAA, 
                          BOSLAACareer>= 0,
                          BOSLAARunsBatted>= 0,
                          BOSLAAGames >= 0,
                          BOSLAARuns >= 0)
              
# Produce a bar chart of the quality rule failures
barplot(BOSLAACheck)

```
Here I see that there is one negative RBI. This would appear to be the only problematic value at this stage.

```{r}
LAACheck <- check_that(LAA, 
                          LAARunsBatted>= 0)

barplot(LAACheck)

BOSCheck <- check_that(BOS, 

                          BOSRunsBatted>= 0)
barplot(BOSCheck)
```
The problem is in Boston. This error could ave various causes.

I can now comfortably look at the data, knowing that I need to watch out for this single negative RBI value. HOpefully the reason for its being will be revealed.

Here ANOVA is very useful. The reason is that it, at a glance, will allow me to work out a lot of things very quickly. one of them is the outliers, and where they are located. I need to be casreful, because I could skew the data in favour of existing assumptions that I have. ANOVA also allows me to look at all of the correlates and get rid of the lower ones and focus on what it important. 


```{r}


```


        
# 2. Exploratory Data Analysis (EDA)

I will do a usmary and histogram of each variable aside from birthdate, and batting hand. This is because they are dates and factorsm which are not suitable for this format.

```{r}
##variablew platerID
##teamID.x
##R - Runs
##G - Games
##H - Hits
##AB - AtBat
##RBI - RunsBatted
##weight 
##height
##salary
##birthDate - Birth
##career.length - Career
##bats
##age
##hit.ind - SeasonHit

summary(BOSLAA)

hist(BOSLAARuns)
lines(density(BOSLAARuns))
hist(BOSLAAGames)
hist(BOSLAAHits)
hist(BOSLAAAtBat)
hist(BOSLAARunsBatted)
hist(BOSLAAWeight)
lines(density(BOSLAAWeight,na.rm=T),col="red",lwd=4)
## for some reaso nwhen I try to plot a line it stays at the bottom of the graph
hist(BOSLAAHeight)
hist(BOSLAACareer)
hist(BOSLAAAge)
hist(BOSLAASeasonHit)
```

## 2.1 EDA plan

*Outline a suitable plan to explore, describe and visualise your data. (5 marks)*  

i will undergo a process including two cyclces which I will switch between at each stage of the data analysis. Most people will use a similar process. This one is inspired by Peng and Matsui's method in their book, the Art of Data Science. They call them epicycles.
The first cycle will be to look at the data, develop assumptions, and then see whether they match.
The second cycle will be ask the questions relevant to these assumptions, see what models might be appropriate, build them, interpret the results, and then report them.

The first epicycle was more or less done in the data cleaning stage. Now I need to build some models.  I will start with some histograms and bar graphs for all data, as well as box plots. Not only this, but means and medians where applicable. I will try to see if any are skewed, and use a Poisson method to measure this.
I will then look at the new data again

ot(BOS, aes(x=BOSBats, y=BOSSalary)) + geom_boxplot()
```

```

## 2.2 EDA and summary of results  

*Undertake and summarise the findings of your data exploration, particularly with respect to the research questions.  Use appropriate summary statistics (uni- and multi-variate) and visualisations. (10 marks)*

I will do a usmary and histogram of each variable

```{r}
hist(BOSCareer)

```

```{r}


```

## 2.3 Additional insights and issues

*Highlight potential further issues or insights uncovered in 2.2.  This might include follow up to findings from your initial EDA.  We accept that the boundary between 2.2 and 2.3 is somewhat arbitrary so use your judgement and maximise good structure and readability. (5 marks)*

```{r}


```

# 3. Modelling

## 3.1 Build a model for player salary

*Given the research question (i.e., the target attribute of salary) outline an analysis plan that incorporates/references any findings from the data cleaning (1.3) and EDA (2.2)  (5 marks). Use R to build a suitable model (10 marks).*  
*NB Submissions where suitable models do not have good fit due to the nature of the data will not be penalised.*  

I want to see the correlates between each variable and player salary. Then, within that, see wether any of those items move win or against one another. Such as career length and age. or runs and bats. This is to see whether I am correct to assume that these elements might be connected.
ANOVA and ANCOVA will be useful for finding a lot of information in one go.
Another thing is that I will go one by one and remove the weakest correlants.

```{r}
step(league)

```

For example, a boc plot will  be useful for the season hit statistic in particular.
##the code below does not work as it should I want two parts. the first being 0, the other being 1. It works for bats, however. Perhaps they need to be changed from intergers to characters?
What I learn here is that batting and salary seem to be related somehow.
For example, a box plot will  be useful for the season hit, and batting hand statistics in particular.

##the code below does not work as it should I want two parts. the first being 0, the other being 1. It works for bats, however. Perhaps they need to be changed from intergers to characters?
What I learn here is that batting and salary seem to be related somehow.

```{r}


ggplot(BOSLAA, aes(x=BOSLAABats, y=BOSLAASalary)) + geom_boxplot()
ggplot(LAA, aes(x=LAABats, y=LAASalary)) + geom_boxplot()
ggplot(BOS, aes(x=BOSBats, y=BOSSalary)) + geom_boxplot()
```
Handedness does not seem to correlate strongly. While ambidextrous seem to earn more than right handed people and even more than left handed people, the latter two have a very large number of outliers.

What about career lenght, and age?
What about birthmonth and age?


```{r}
##testing correlations is important here
cor(BOSLAAAge, BOSLAASalary, method = c("pearson", "kendall", "spearman"))
cor.test(BOSLAAAge, BOSLAASalary, method=c("pearson", "kendall", "spearman"))
```
## 3.2 Critique model using relevant diagnostics

*Offer an interpretation of the model characteristics, goodness of fit and graphical diagnostics (5 marks) for the model built in 3.1. Explain any potential weaknesses (5 marks).*

In order to test this model, I will want to see what happens with other teams. How accurate is it then?
Do I have overfitting?

```{r}


```


## 3.3 Suggest improvements to your model

*Based on the findings in 3.2 articulates possible alternative approaches to address them (5 marks).*
The easiest and most obvious way to improve the model would be to have even more data to compare to. This will allow one to track chenges over time. Player salary could behave differently according to the player profile over time. Some players would have high salaries over a short period, while others could have high salaries over a long period.
Luckily, I do have more data. I can compare it with the league as a whole. Not only this, but I could also get a hold of data over time.
```{r}


```


# 4. Extension work

## 4.1 Model the likelihood of a player having scored a Hit (using the hit.ind variable provided).

*Given this second research question (i.e., involving the binary target attribute) provide a plan of analysis based on relevant EDA for this attribute (10 marks). The model is described, explained and critiqued (10 marks).*
*NB Submissions where suitable models do not have good fit due to the nature of the data will not be penalised.* 

This will be done on four levels. The whole set, two teams, and then each of the two teams. This iwll be very similar to how the model for player salary was developed.

```{r}
table(BOSLAASeasonHit)
mean(BOSLAASeasonHit)

```
```{r}
CrossTable(BOSLAASeasonHit, BOSLAABats, digits=2, prop.r = TRUE, prop.c = TRUE,  prop.chisq = FALSE, 
           chisq = TRUE, fisher = TRUE)
```
##crosstable shows more information
# References  

*Add any references here. NB You can either do this manually or automatically with a `.bib` file (which then must be submitted along with your `.Rmd` file).  See the RMarkdown [documentation](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html) for guidance.*    
