```{r}
sample = dbinom(0:1000, size = 1000, prob = 0.5)

pbinom(0.5, 1)


```

The binomial distribution is a discrete probability distribution. It describes the outcome of n independent trials in an experiment. Each trial is assumed to have only two outcomes, either success or failure. If the probability of a successful trial is p, then the probability of having x successful outcomes in an experiment of n independent trials is as follows.

let's go


```{r}
nsamples = 10000
s_norm = 0
s_std = 1
y = rnorm(nsamples, s_norm, s_std)
y2 = rbinom(nsamples, 1000, 0.5)

plot(y)
plot(y2)

```
Density, distribution function, quantile function and random generation for the Poisson distribution with parameter lambda.

Usage
dpois(x, lambda, log = FALSE)
ppois(q, lambda, lower.tail = TRUE, log.p = FALSE)
qpois(p, lambda, lower.tail = TRUE, log.p = FALSE)
rpois(n, lambda)
```{r}
hist(y)
hist(y2)

```
Density, distribution function, quantile function and random generation for the normal distribution with mean equal to mean and standard deviation equal to sd.

Usage
dnorm(x, mean = 0, sd = 1, log = FALSE)
pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
rnorm(n, mean = 0, sd = 1)
```{r}
seq(0, 5, by=0.3)

```

seq(from, to)
seq(from, to, by= )
seq(from, to, length.out= )
seq(along.with= )
seq(from)
seq(length.out= )


```{r}
c<-seq(60,150)
d<-dnorm(c,mean = 105, sd=9)

plot(c, dnorm(c,mean = 105, sd=9), type = "l")

meang = 105
stdg = 9

```
https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/plot


```{r}

plot(c, dnorm(c,mean = 105, sd=9), type = "l")
abline(v=90, lty=2)
abline(v=125, lty=3)

```

ggplot also does the below
```{r}

dnorm(meang, mean=meang, sd = stdg)
pnorm(meang, mean=meang, sd = stdg)
qnorm(0.1, mean = meang, sd = stdg)

```


part2
```{r}
mpg= auto_mpg

summary(mpg)
````

part 2


```{r}

ggplot(mpg)+aes(y=acceleration, x=horsepower) + geom_point()

````



```{r}
ggplot(data=mpg, aes(x=acceleration)) + geom_density() +
theme_light() + ggtitle("Histogram of Acceletation")

```

```{r}

plot(mpg$horsepower, mpg$acceleration, type="l")
````

Compute a 95% confidence interval for the mean for one of the variable that appears to be normally
distributed

```{r}

qnorm(0.975)

mean.acceleration<-mean(mpg$acceleration)
sd.acceleration<-sd(mpg$acceleration)


ucl<-mean.acceleration-qnorm(0.975)*sd.acceleration/sqrt(398)
lcl<-mean.acceleration+qnorm(0.975)*sd.acceleration/sqrt(398)

ucl

lcl


```
ucl is  confidewnce limit
lcl is lower confidence limite

```{r}

hist(mpg$acceleration)

````
      Compute a 95% confidence interval for the proportion of cars from the 80s?

```{r}

table(mpg$`model year`)
sum(mpg$`model year`>75)

prop.80<-sum(mpg$`model year`>79)/nrow(mpg)

vr<-prop.80*(1-prop.80)/nrow(mpg)
prop.ucl<-prop.80+ qnorm(0.975)*sqrt(vr)
prop.lcl<-prop.80- qnorm(0.975)*sqrt(vr)

prop.lcl
prop.ucl

````


ok

      